### 2010-12-28 Documentation/atropos/atropos.tex
ENH:  A few minor edits.
git-svn-id: https://advants.svn.sourceforge.net/svnroot/advants@728 fd2da0a4-d968-48df-9cb5-d8c496287cbf

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
                           [initialMeshResolution,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output [correctedImage,<biasField>]
     -h 
     --help 
\end{lstlisting} 
\end{command}
\subsubsection{Integration of N4 Bias Correction}
The typical segmentation processing pipeline begins with an intensity
normalization/bias correction step using the recently developed N4
algorithm \citep{Tustison2010} which is based on the popular
non-parametric nonuniformity intensity normalization (N3) introduced in \cite{Sled1998}.  It was previously mentioned that
several methods proposed in the literature have taken an integrative
view of the segmentation problem by incorporating an intrinsic bias
correction step into the actual workflow.  The
advancements introduced with N4 permit such an adaptive
integration with Atropos.  Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 \citep{Boyes2008}.  Thus, when performing 3-tissue segmentation, we
may opt to use, for instance, the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.  This is done by setting
the `{\ttfamily --weight-image}' option on the N4 command line call
(see Listing \ref{listing:n4}) to the posterior probability image
corresponding to the white matter produced as output in the Atropos
call, i.e. `{\ttfamily Atropos --output}'.  N4 is
currently being added as a standard part of the Insight Toolkit.  The
evaluation section will illustrate inclusion of Atropos, N4 and ANTs
in a brain processing pipeline.  An example of this procedure, using brainweb data with
40\% RF bias, is in figure~\ref{fig:bwebrf40}.  We supply the
information necessary to repeat this example in the script entitled
`{\ttfamily  atroposBwebRF40FigureExample.sh}' which is available in the ANTs toolkit's Atropos
documentation folder as of svn commit 711.
\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6in]{Figures/bwebN4ex.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small We combine N4 and Atropos by simple
sequential processing and apply to brainweb single-subject data with
40\% RF bias and 3\% noise.  The $\beta$ for the MRF term is, here,
$0.2$.  Slice 71 of the input data is in (a).  The initial K-means
($K=3$) segmentation is in panel (b).  We use the brain mask to guide
N4 bias correction and produce the image, shown in (c).  We repeat the
K-means segmentation, but with the N4 corrected image as input, and
produce the segmentation in (d).  The average 3-tissue Dice overlap of
result (b) is 0.906 while the average overlap for (d) is 0.954.
Arrows highlight a region of large before-after segmentation
discrepancy.}
\label{fig:bwebrf40}
\end{center}
\end{figure}

@@

\url{http://mouldy.bic.mni.mcgill.ca/brainweb/}.  We employ both the

@@

\subsubsection{20 Subject Evaluation} The single-subject brainweb

@@

subject brainweb data allows us to use 2-fold cross-validation to test

@@

\subsection{Reproducibility of this Evaluation}
The brainweb data is freely available.  We used single-subject
brainweb data as is but added a meta format data header to the raw
binary files.   An example copy of this header is contained in {\verb atroposBwebRF40FigureExample.sh }.  The 20 subject data, however, required excluding
non-cerebrum tissue classes.  The Hammers data was also used as is.
The scripts for the single-subject brainweb study is available in
supplementary material.  The initialization procedure for the brainweb 20 and
the Hammers evaluation data is based on freely available scripts
included with the ANTs toolkit, specifically  {\verb ants_multitemplate_labeling.sh } and  {\verb atroposBwebRF40FigureExample.sh }.

\section{Discussion}
We introduced Atropos, the theory and implementation
details and documented its performance in a variety of use cases.  We
also showed evidence that the openly available N4 bias correction can easily used with Atropos to improve segmentation.
Furthermore, we used multiple subject brainweb data to build dataset
specific priors that provided the most consistent segmentation
performance across tissues.  Finally, we used majority voting to
initialize an Atropos EM solution to a 69 class brain parcellation
problem.  Significant improvements were gained in multiple brain
regions, in particular in temporal lobe cortex, the hippocampi and
amygdalae and the lateral ventricles.  This work, in summary, proves
the applicability of Atropos in both basic and extended use cases.

\subsection{Performance on BrainWeb Data}
Atropos results are competitive with the state of the art.  For
instance, \cite{Ashburner2005} (SPM5) evaluated on 0\% RF (bias field) 3\%
noise brainweb single subject data finding 0.932 (GM) and .961 (WM)
Dice. Results on 40\% RF were 0.934 (GM) and 0.961 (WM). SPM5 exhibits
insensitivity to bias similar to our own best results on the 40\%RF, 3\% 
noise case (MRF-$\beta$=0.2, K-means $+$ N4) with Dice for GM is
0.950737 and for WM is 0.962586.  \cite{Nakamura2009} gave GM Dice
results (brainweb single 3\% noise) of 0.962 (0\%RF), 0.964(20\%RF)
and 0.956(40\%RF) which are slightly higher than either SPM5 or Atropos
results.  However, \cite{Nakamura2009} do not report WM or CSF numbers
for comparison.  Topology preserving methods also perform well.
\cite{Shiee2010} achieved Dice results for 3\% noise 20\% RF brainweb
single subject with 0.927, 0.912 and 0.900 for WM, GM and CSF,
respectively.  These are excellent numbers given the additional
constraint applied to the segmentation.  
\cite{Bazin2007c} proposed TOADS and, estimating from the
paper's graph, showed that the average Dice accuracy for 3\% noise various RF was
0.95-0.96 for WM, 0.93-0.95 for GM and 0.92-0.94 for CSF.
Perhaps the most recent balanced evaluation was performed in
\citep{Klauschen2009}, which reports confusion matrix numbers, rather
than Dice overlap.  Because the absolute true number of GM and WM
voxels for brainweb are known, we can convert the confusion matrix to
Dice.  In that case, the SPM5 Dice for brainweb single GM and WM is 0.8847
and 0.9085, respectively, while FreeSurfer and FSL's accuracy is
lower. The best GM result for 20 subjects is obtained by SPM5: 0.9295;
best WM Dice is from FSL: 0.9499.  We note that \cite{Klauschen2009}
used a comprehensive evaluation where quality of brain extraction also
contributed to the outcome.  Thus, the results must be interpreted
slightly differently than those from other papers.

@@

\subsection{Clinically-Related Evaluation}
While specifying performance on brainweb is highly
valuable, clinical validation is a second important aspect of
segmentation evaluation.  For instance, 
\citep{Freeborough1997,Westlye2009,Sanchez-Benavides2010,Chou2009,Bresser2011}
are only a few of the papers that evaluate segmentation performance with
respect to a known neurobiological outcome measure.  Atropos is
currently used in clinical studies and a number of clinically-focused,
application-specific evaluations are ongoing and will constitute
future work.  One early example of a clinically-focused Atropos
neuroimaging application is in \citep{Avants2010c}.  An additional successful
application area is that of ventilation-based segmentation of hyperpolarized helium-3 MRI
\citep{Tustison2010a} which also used the open-source  Glamorous Glue
algorithm to impose topology constraints.  Thus, future work may
incorporate topology more closely into the Atropos methodology.

>>>>

OPTIONS: 
     -d, --image-dimensionality 2/3/4
     -i, --input-image inputImageFilename
     -x, --mask-image maskImageFilename
     -w, --weight-image weightImageFilename
     -s, --shrink-factor 1/2/3/4/...
     -c, --convergence [<numberOfIterations=50>,<convergenceThreshold=0.001>]
     -b, --bspline-fitting [splineDistance,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
                           [initialMeshResolution,<splineOrder=3>,<sigmoidAlpha=0.0>,
                           <sigmoidBeta=0.5>]
     -t, --histogram-sharpening [<FWHM=0.15>,<wienerNoise=0.01>,<numberOfHistogramBins=200>]
     -o, --output [correctedImage,<biasField>]
     -h 
     --help 
\end{lstlisting} 
\end{command}
\subsubsection{Integration of N4 Bias Correction}
The typical segmentation processing pipeline begins with an intensity
normalization/bias correction step using the recently developed N4
algorithm \citep{Tustison2010} which is based on the popular
non-parametric nonuniformity intensity normalization (N3) introduced in \cite{Sled1998}.  It was previously mentioned that
several methods proposed in the literature have taken an integrative
view of the segmentation problem by incorporating an intrinsic bias
correction step into the actual workflow.  The
advancements introduced with N4 permit such an adaptive
integration with Atropos.  Recent demonstrations suggest improved
white matter segmentation produces better gain field estimates using
N3 \citep{Boyes2008}.  Thus, when performing 3-tissue segmentation, we
may opt to use, for instance, the posterior probability map of white matter at the current
iteration as a weighted mask for input to N4.  This is done by setting
the `{\ttfamily --weight-image}' option on the N4 command line call
(see Listing \ref{listing:n4}) to the posterior probability image
corresponding to the white matter produced as output in the Atropos
call, i.e. `{\ttfamily Atropos --output}'.  N4 is
currently being added as a standard part of the Insight Toolkit.  The
evaluation section will illustrate inclusion of Atropos, N4 and ANTs
in a brain processing pipeline.  An example of this procedure, using BrainWeb data with
40\% RF bias, is in figure~\ref{fig:bwebrf40}.  We supply the
information necessary to repeat this example in the script entitled
`{\ttfamily  atroposBwebRF40FigureExample.sh}' which is available in the ANTs toolkit's Atropos
documentation folder as of svn commit 711.
\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6in]{Figures/bwebN4ex.pdf}
\end{tabular}
\caption{\baselineskip 12pt \small We combine N4 and Atropos by simple
sequential processing and apply to BrainWeb single-subject data with
40\% RF bias and 3\% noise.  The $\beta$ for the MRF term is, here,
$0.2$.  Slice 71 of the input data is in (a).  The initial K-means
($K=3$) segmentation is in panel (b).  We use the brain mask to guide
N4 bias correction and produce the image, shown in (c).  We repeat the
K-means segmentation, but with the N4 corrected image as input, and
produce the segmentation in (d).  The average 3-tissue Dice overlap of
result (b) is 0.906 while the average overlap for (d) is 0.954.
Arrows highlight a region of large before-after segmentation
discrepancy.}
\label{fig:bwebrf40}
\end{center}
\end{figure}

@@

\url{http://mouldy.bic.mni.mcgill.ca/BrainWeb/}.  We employ both the

@@

\subsubsection{20 Subject Evaluation} The single-subject BrainWeb

@@

subject BrainWeb data allows us to use 2-fold cross-validation to test

@@

\subsection{Reproducibility of this Evaluation}
The BrainWeb data is freely available.  We used single-subject
BrainWeb data as is but added a meta format data header to the raw
binary files.   An example copy of this header is contained in {\verb atroposBwebRF40FigureExample.sh }.  The 20 subject data, however, required excluding
non-cerebrum tissue classes.  The Hammers data was also used as is.
The scripts for the single-subject BrainWeb study is available in
supplementary material.  The initialization procedure for the BrainWeb 20 and
the Hammers evaluation data is based on freely available scripts
included with the ANTs toolkit, specifically  {\verb ants_multitemplate_labeling.sh } and  {\verb atroposBwebRF40FigureExample.sh }.

\section{Discussion}
We introduced Atropos, the theory and implementation
details and documented its performance in a variety of use cases.  We
also showed evidence that the openly available N4 bias correction can easily used with Atropos to improve segmentation.
Furthermore, we used multiple subject BrainWeb data to build dataset
specific priors that provided the most consistent segmentation
performance across tissues.  Finally, we used majority voting to
initialize an Atropos EM solution to a 69 class brain parcellation
problem.  Significant improvements were gained in multiple brain
regions, in particular in temporal lobe cortex, the hippocampi and
amygdalae and the lateral ventricles.  This work, in summary, proves
the applicability of Atropos in both basic and extended use cases.

\subsection{Performance on BrainWeb Data}
Atropos results are competitive with the state of the art.  For
instance, \cite{Ashburner2005} (SPM5) evaluated on 0\% RF (bias field) 3\%
noise BrainWeb single subject data finding 0.932 (GM) and .961 (WM)
Dice. Results on 40\% RF were 0.934 (GM) and 0.961 (WM). SPM5 exhibits
insensitivity to bias similar to our own best results on the 40\% RF, 3\% 
noise case (MRF-$\beta$=0.2, K-means $+$ N4) with Dice for GM is
0.950737 and for WM is 0.962586.  \cite{Nakamura2009} gave GM Dice
results (BrainWeb single 3\% noise) of 0.962 (0\%cRF), 0.964 (20\% RF)
and 0.956 (40\%RF) which are slightly higher than either SPM5 or Atropos
results.  However, \cite{Nakamura2009} do not report WM or CSF numbers
for comparison.  Topology preserving methods also perform well.
\cite{Shiee2010} achieved Dice results for 3\% noise 20\% RF BrainWeb
single subject with 0.927, 0.912 and 0.900 for WM, GM and CSF,
respectively.  These are excellent numbers given the additional
constraint applied to the segmentation.  
\cite{Bazin2007c} proposed TOADS and, estimating from the
paper's graph, showed that the average Dice accuracy for 3\% noise various RF was
0.95--0.96 for WM, 0.93--0.95 for GM and 0.92--0.94 for CSF.
Perhaps the most recent balanced evaluation was performed in
\citep{Klauschen2009}, which reports confusion matrix numbers, rather
than Dice overlap.  Because the absolute true number of GM and WM
voxels for BrainWeb are known, we can convert the confusion matrix to
Dice.  In that case, the SPM5 Dice for BrainWeb single GM and WM is 0.8847
and 0.9085, respectively, while FreeSurfer and FSL's accuracy is
lower. The best GM result for 20 subjects is obtained by SPM5: 0.9295;
best WM Dice is from FSL: 0.9499.  We note that \cite{Klauschen2009}
used a comprehensive evaluation where quality of brain extraction also
contributed to the outcome.  Thus, the results must be interpreted
slightly differently than those from other papers.

@@

\subsection{Clinically-Related Evaluation}
While specifying performance on BrainWeb is highly
valuable, clinical validation is a second important aspect of
segmentation evaluation.  For instance, 
\citep{Freeborough1997,Westlye2009,Sanchez-Benavides2010,Chou2009,Bresser2011}
are only a few of the papers that evaluate segmentation performance with
respect to a known neurobiological outcome measure.  Atropos is
currently used in clinical studies and a number of clinically-focused,
application-specific evaluations are ongoing and will constitute
future work.  One early example of a clinically-focused Atropos
neuroimaging application is in \citep{Avants2010c}.  An additional successful
application area is that of ventilation-based segmentation of hyperpolarized helium-3 MRI
\citep{Tustison2010a} which also used the open source Glamorous Glue
algorithm to impose topology constraints \cite{Tustison2010b}.  Thus, future work may
incorporate topology more closely into the Atropos methodology.
